Long Range Interactions 3.
========================================================

### Data Plotting

The plotting of this data is now quite simple although the statistical reasoning behind comparisons remains quite subtle - given the fact that the sites have different frequencies - althought this shouldn't effect the plots??? Needs thought.

```{r plotLRI, message=FALSE}
# I am now working on the iMac
load("res.RData")
ls()
head(res1Chr1)

library(ggplot2)
qplot(x=log10(abs(start1-start2)), y=observed/expected, data=res1Chr1, facets=site1~site2, geom='smooth')+theme_bw()
```

***This is not the relation plotted by Wen-Ching!!! - It's just an example.***

More work needed on this:

1. Need to check through Wen-Ching script to see exactly the expected-observed relation he is plotting.
2. First check all (or a reasonable selection) of chromosomes to check they all behave similarly.
3. If all similar then aggregate the data - maybe parallelising the code to use multiple cores???
4. Maybe interested in self-self bin interactions???
5. Create simulated data to check correctness of code!!!


est. 2-3 days work.

6. Install RSQLite on CS server so analysis can be run there.
7. Create package, scripts and man pages so group can run these analyses.

est... dunno ???


### More Work

I've found it a little difficult to follow Wen-Chings code. I'm concerned I have quite different curves - all very similar. The best I can work out he is looking at log10 distance vs the log2(observed/expected). Where this produces -Inf values (log of zero) he replaces with a small positive value (0.5 in his code if I understand correctly --  0.001 in mine below)

```{r moreWork, message=FALSE}
temp = with(res1Chr1, log2(observed/expected))
temp[temp == -Inf]<-0.001
res1Chr1$temp=temp
qplot(x=log10(abs(start1-start2)), y=temp, data=res1Chr1, facets=site1~site2, geom='smooth')+theme_bw()

# all on the same plot
qplot(x=log10(abs(start1-start2)), y=temp, data=res1Chr1, colour=paste(site1,site2), geom='smooth')+theme_bw()

# bin2d
qplot(x=log10(abs(start1-start2)), y=temp, data=res1Chr1, facets=site1~site2, geom='blank')+theme_bw()+geom_bin2d()


```
